% !TeX spellcheck = en_US
\chapter{Conclusion}%
%WAS MUSS NOCH GEMACHR WERDEN DASSS ES Auf echtem Gcode und \acrshort{CAM} l채uft ???
\section{Summary}%
This thesis proposes a method for optimizing the execution of a toolpath on manufacturing machines with redundant \acrshort{DoF}s. Firstly, the problem formulation (Chapter \ref{Problem Formulation}) highlights the flexibility and various issues that arise with redundant \acrshort{DoF}s. It discusses how singularity avoidance can help industrial robots avoid sub-optimal poses that may lead to unexpected behavior. Redundant \acrshort{DoF} can also affect joint accelerations and jerk, potentially causing excessive wear on parts and resulting in more downtime and sub-optimal part quality. Additionally, other factors such as extension control, precision, and energy usage are mentioned, and their impact on the manufacturing process is discussed.

The aim of this thesis (Chapter \ref{OBJECTIVE}) is to propose and validate a systematic approach to leveraging these redundant \acrshort{DoF}s in order to optimize towards a user-defined goal. Currently, there is no publication available that presents a general solution to this problem.


Chapter \ref{SST} provides a comprehensive discussion on the state of science and technology, aiming to provide a clear understanding of the individual components of manufacturing systems and optimization algorithms. Both subtractive and additive manufacturing are examined in detail, including a review of their respective strengths and weaknesses. An important focus is placed on one of the common processes in additive manufacturing, namely \acrshort{WAAM}, which has a significant relationship with industrial robots. In Chapter \ref{IR}, the functionality of these robots is described in detail, with special attention given to the issue of redundancy in such robotic systems.

Chapter \ref{comp} focuses on a comparative analysis of published research papers related to singularity avoidance, optimization of joint acceleration and jerk, optimization of energy use, and optimization of stiffness. Each section highlights the available options and approaches for leveraging redundant \acrshort{DoF}s to achieve improved performance in these specific areas. The examination of multiple methods serves as the foundation for understanding the current state of cutting-edge research and identifying any existing research gaps.
Thus far, no global optimization approach has been proposed that can consider a user-defined input with specified goals, based on multiple process variables, and providing the optimal settings for the redundant \acrshort{DoF}s. This represents an important research opportunity in the field.

Chapter \ref{METmain} presents a solution for the identified research gap. In order to address this problem, a carefully selected list of process variables is introduced and thoroughly discussed. These variables can derived from the movement of a robots' arm and includes variables such as the rotational position of joints over time, as well as their subsequent derivatives and direction changes.
After summarizing which process variables can be extracted form a toolpath traversed by a industrial robot, the two main parts of the method are validated. 


The validation process involves constructing a mathematical model that represents an industrial robot. Additionally, three artificial toolpaths are generated for the robot model to traverse. These toolpaths are defined in five \acrshort{DoF}s, while the 6th \acrshort{DoF}, rotation around the Z-axis, is not defined and can be chosen freely. The range of rotation is set from -135째 to +135째, divided into 5째 increments. For each possible setting of the rotation around the Z-axis, the inverse kinematics algorithm calculates the joint positions over time required to traverse the toolpaths. Subsequently, the process variables are extracted and compared. Based on their numerical values and user-defined importance factors, the global score for each setting is calculated. The same process is then performed on a lager production-grate toolpath. This approach allows for the analysis of different settings for the redundant \acrshort{DoF} and the identification of the most optimal one.

The same process is used to analyze the results when two different redundant \acrshort{DoF}s are present. The first redundancy is in the form of the rotation around the Z-axis, and the second one is introduced as the tilting of a rotary-tilt table. In this scenario, the goal is to analyze how each combination of the two redundant \acrshort{DoF}s affects the analyzed process variables. The resulting scores are shown in the form of a matrix. The matrices for the three different toolpaths exhibit different characteristics but generally result in a fairly smooth surface.

The second step in the presented method involves the optimization of the redundant \acrshort{DoF}. Instead of calculating the joint positions for every possible setting, a \acrshort{PSO} algorithm is utilized. This implementation saves a significant amount of time and has the capability to quickly converge to the global optima. The population size is set to 20, and the iterations are capped at five.
As mentioned in the discussion of the results, the performance of the PSO algorithm depends on the initial positions of the individual particles.

One of the proposed goals mentioned in the task for this thesis is the implementation of the optimization loop in \acrshort{CAM}. This implementation would allow for the validation of the described optimization process in Chapter \ref{CAMinloop}. However, due to the limited time-frame of the thesis and the unavailability of a developer version of the selected \acrshort{CAM} software with the necessary \acrshort{API}, this goal could not be achieved.


\section{Outlook}\label{Outlook}

This work focuses only on on a select number of process variables, but this selection can be expanded to further optimize the real-world manufacturing process. One potential additional variable that can be optimized, is the stiffness value of the robotic system based on its current pose. Maintaining high stiffness in the orthogonal cutting direction is crucial for minimizing deviations. Low stiffness combined with high contact forces can result in significant dimensional errors. Finite-element analysis or multi-body simulation can be used to determine the stiffness value and optimize it for better performance by specifically constraining the redundant \acrshort{DoF}.

Another important variable for future research is the collision index. The collision index is used to identify potential collisions between any part of the robot or the end-effector and the workpieces or other objects in the environment. This variable is particularly significant in scenarios involving \acrshort{WAAM} systems, where loose wires can change their position depending on the robots pose. Preventing collisions is essential to avoid damage to the workpiece, the robot, or other equipment. As of now the optimization algorithm does not consider this option.

One area where the \acrshort{PSO} optimization process can lead to significant advantages is section-wise optimization. The toolpath can be divided into sections, and for each section, the optimal settings of the redundant \acrshort{DoF} can be determined. To make this approach work effectively, it is important to consider the boundary conditions at the transition points between the individual sections.
Furthermore, implementing the proposed method in a \acrshort{CAM} software can provide faster computation and enable the optimization of more complex toolpaths. Another option to speed up the computation is to program the algorithm in such a way that it is designed for multi-threading. Such implementation can result in significantly faster optimization.  

Moving forward, it is necessary to conduct further validation processes, including real-world tests and simulations, to evaluate the performance and verify the effectiveness of the proposed methodology. Additionally, exploring the implementation of \acrshort{ML} techniques for best-case calculations could be a valuable avenue for future research.

%In conclusion, conducting additional tests and thorough validation can greatly enhance the efficiency, productivity, and safety of the manufacturing process.

%Energy Use:
%Redundant degrees of freedom can increase energy consumption in manufacturing systems. Additional joints and actuators require more power, leading to higher energy costs and environmental impact. Efficient energy management strategies, such as optimizing control algorithms, utilizing regenerative braking, and incorporating energy-efficient components, are crucial to minimize energy use while maintaining system performance.