% !TeX spellcheck = en_US
\chapter{Introduction}%
\section{Motivation}%

In the age of "Industrie 4.0", advanced technologies like digital twins, have greatly transformed industrial manufacturing~\cite{Singh.2021}. A considerable amount of data can be gathered from various processes, like milling or 3D printing. By analyzing this data, it is possible to find new and optimized methods for improving efficiency as well as streamline and enhancing the manufacturing process~\cite{Ghobakhloo.2020}. By doing so, a significant amount of resources, like time and money, can be saved while at the same time increasing the quality of the produced product~\cite{Bibby.2018,Simonis.2016}.
Computer-aided manufacturing (\acrshort{CAM}) has been introduced as a crucial tool to improve productivity and accuracy in creating customized products~\cite{Feldhausen.2022}. \acrshort{CAM} systems automate and optimize tasks such as machining, welding, and assembly~\cite{LalitNarayan.2013b}. One of the key strengths of \acrshort{CAM} lies in its precision and consistency, ensuring that intricate components are produced with minimal error. Furthermore, \acrshort{CAM} systems contribute to increased efficiency by minimizing material waste and reducing production time~\cite{Dubovska.2014}. These capabilities play a significant role in achieving a carbon-neutral production process~\cite{Saxena.2020}. One of the most important areas of \acrshort{CAM} is the calculation of the tool path for computer numerical control (\acrshort{CNC}) machines as well as the movement and behavior of multi-axis industrial robots~\cite{Pan}. 


Manufacturing machines are the backbone of modern industrial processes~\cite{Bi.2020}. These machines encompass a wide range of equipment, from \acrshort{CNC} machining centers to 3D printers and automated assembly lines. Their primary ability lies in precision  and efficiency. \acrshort{CNC} machines, for instance, can repeatedly produce intricate parts with high accuracy ($\mu$m-scale), reducing human error and ensuring consistency~\cite{Jia.2018,Liberman.2021}. Industrial robots are a dominant part in the area of manufacturing as they can perform multi-axis movements. These capabilities are especially helpful to fulfill the customers wishes for individualized products and one-off parts~\cite{Sherwani.2020}. They are cheaper to acquire and easily adapted for different tasks compared to \acrshort{CNC} milling machines, but have their own set of disadvantages like for example lower stiffness~\cite{Iglesias.2015, Liberman.2021}. One of the most important advantages of industrial robots is their wide adaptability. They allow for quick reconfiguration to produce different components or products, promoting flexibility in manufacturing~\cite{Billard.2019}. Further, advancements in robotics and Artificial Intelligence (\acrshort{AI}) have broadened their capabilities, enabling tasks that were once deemed too complex or hazardous for humans~\cite{Goel.2020}. 

Achieving better efficiency and continuous sustainability in the current fast-changing environment, requires a thorough analysis of the interdependent relationships between the manufactured part, the user-defined process parameters, and boundary conditions that govern multi-axis robot programs~\cite{Pan, Gadaleta.2019}. Companies that work with industrial robots can place a strong emphasis on energy reduction, cycle-time minimization, or precision. By defining optimal boundary conditions, it is possible to optimize these process variables. \acrshort{CAM} enables the simulation of the planned process, thus adapting any boundary conditions in advance, to fit the selected goals~\cite{Kyratsis.2020,Maiti.2017,Pan,Uhlmann.2016}.
This thesis is focused on a methodical approach for analyzing process variables and optimizing boundary conditions in multi-axis robot programs. 

\section{Problem Formulation}\label{Problem Formulation}
Manufacturing systems that incorporate redundant degrees of freedom (\acrshort{DoF}s) offer significant advantages in terms of flexibility and adaptability~\cite{Anjum.2022}. One example of a system with redundancy is a 6-\acrshort{DoF} industrial robot with a rotary tilt table, which brings the system to eight \acrshort{DoF}s. However, these systems also present various conflict points that need to be carefully managed to ensure optimal performance~\cite{Boscariol.2020}.


One of the critical challenges in manufacturing systems with redundant \acrshort{DoF}s is singularity avoidance~\cite{Kireanski.1993, Wang.2022}. Singularities, which are critical points in the motion of a robot manipulator, arise when the system loses its ability to maintain full control or achieves limited mobility as a consequence of specific joint configurations~\cite{Malyshev.2022}. These configurations result in the loss of a \acrshort{DoF} or make the system highly sensitive to small changes, leading to unstable or even unpredictable and dangerous behavior~\cite{Zhao.2021, Milenkovic.2021}. Limiting the possible positions by adding artificial constraints can help to avoid this problem~\cite{Faria.2018}. %Advanced control algorithms and motion planning techniques are necessary to detect and avoid singularities, ensuring the system operates within safe and stable regions.

%Redundant degrees of freedom introduce complexities in controlling joint accelerations and jerks. Rapid changes in acceleration and jerk can cause mechanical stress, decrease system lifespan, and compromise precision. Proper trajectory planning and control schemes that consider joint acceleration and jerk limits are crucial to maintain smooth and controlled motion, reducing the risk of mechanical failures and ensuring accurate positioning.

%Extension:
%Redundant degrees of freedom can provide additional extension capabilities to industrial robots, allowing them to reach difficult-to-access areas. However, managing and controlling the extension can be challenging, particularly when precise positioning or maintaining stability is required. Accurate modeling, control algorithms, and calibration techniques are necessary to ensure proper extension control and to avoid issues related to overextension or underextension.

%Energy Use:
%Redundant degrees of freedom can increase energy consumption in manufacturing systems. Additional joints and actuators require more power, leading to higher energy costs and environmental impact. Efficient energy management strategies, such as optimizing control algorithms, utilizing regenerative braking, and incorporating energy-efficient components, are crucial to minimize energy use while maintaining system performance.

%Direction Changes:
%Redundant degrees of freedom offer increased flexibility in changing the direction of motion. However, abrupt and frequent direction changes can lead to mechanical stress, decreased precision, and increased energy consumption. Proper trajectory planning, smooth path generation, and optimized control algorithms are essential to manage direction changes effectively, ensuring smooth and controlled motion without compromising system performance.

%Transfer Time:
%Transfer time is another critical aspect affected by redundant degrees of freedom. Complex motion patterns required for utilizing redundant degrees of freedom can increase the time required to complete a task. Minimizing transfer time is crucial for improving production efficiency and throughput. Optimal path planning, motion optimization, and parallel processing techniques can be employed to reduce transfer time while leveraging redundant degrees of freedom effectively.

%Precision:
%Redundant degrees of freedom can enhance precision and accuracy in manufacturing systems. However, achieving and maintaining high precision can be challenging due to increased complexity and sensitivity to various factors. Calibration techniques, accurate modeling, feedback control, and error compensation algorithms are vital to ensure precise positioning and improve overall system accuracy.

%Maximum Load Capacity:
%Redundant degrees of freedom can influence the maximum load capacity of industrial robots. Increased complexity and additional joints can affect the robot's ability to handle heavy loads or maintain stability. Proper design considerations, such as structural reinforcement, joint torque and force limitations, and payload distribution, are necessary to ensure safe and reliable operation under maximum load conditions.

%Stiffness:
%The presence of redundant degrees of freedom can affect the stiffness of the manufacturing system. Increased flexibility due to redundant degrees of freedom may introduce compliance and reduce overall system stiffness. This can impact precision, accuracy, and stability. Careful design, utilization of appropriate materials, and control strategies that compensate for compliance can be employed to ensure the desired level of stiffness and system rigidity.


%---------------





Another significant aspect of manufacturing systems with redundant \acrshort{DoF}s is joint acceleration and jerk, which is the rate of change of acceleration. The robot must allocate accelerations effectively among its joints to achieve smooth and coordinated motion. Failure to do so can result in jerky or erratic movements, which not only compromises precision but also impacts the efficiency of the manufacturing process~\cite{Duong.2021}. Rapid changes in acceleration and jerk can cause mechanical stress and thus, decrease the system lifespan. Additionally, the joints can be limited in their ability to keep up with the required speed due to limitations in power~\cite{Staff.1988}. Therefore, advanced control algorithms and motion planning techniques are necessary to optimize joint motion and minimize conflicts in joint acceleration and jerk~\cite{Duong.2021, Valente.2017}. As of now, no publication is considering the option of constraining the redundant \acrshort{DoF}s, to effectively mitigate the occurring joint accelerations while at the same time considering further process variables. 

Extension control, which is a further process variables, is a critical aspect that needs to be addressed in systems with redundant \acrshort{DoF}s. Redundant \acrshort{DoF}s can provide additional extension capabilities to industrial robots, allowing them to reach difficult-to-access areas~\cite{Duong.2021}. However, managing and controlling the extension can be challenging, particularly when precise positioning or maintaining stiffness is required~\cite{Lin.2022}.
The robot must accurately determine the appropriate position for each joint to avoid unnecessary over-extension and collisions with the surrounding environment. The robot pose, which is the combination of position and orientation in three-dimensional space, also has a significant effect on robot stiffness~\cite{Xiong.2019}. An increased number of joints can introduce more play and reduce overall system stiffness. This can affect precision, accuracy and stability. Robot pose and its \acrshort{DoF}s must be carefully considered to ensure the desired level of system rigidity~\cite{Wang.2022, Shi.2021}. The current state of the art does not provide a methodical approach that, can utilize the redundant \acrshort{DoF}s to ensure a most optimal solution for multiple process variables like, robot stiffness or joint movements.


Precision is a crucial element in manufacturing systems, and closely tied to its stiffness. The robot needs to have precise control over the movement of each joint to achieve the desired accuracy in the manufacturing process. However, achieving and maintaining high accuracy and repeatability can be difficult due to the increased complexity~\cite{Duong.2021}. % and sensitivity to various factors %Any deviations or errors in joint motion can lead to compromised precision and reduced product quality. %To address this, advanced sensing and feedback control mechanisms, such as force/torque sensors and visual feedback systems, can be employed to enhance precision and ensure the accurate execution of manufacturing tasks.
Frequent changes in direction in the joints are another factor that affects precision. 
Due to the serial kinematics of industrial robots, the present play in the motor joints can add up the inaccuracies and impede the manufacturing process~\cite{Huynh.2020, ChenGang.2014}. Mechanical stresses, decreased precision, and increased energy consumption can all result from abrupt and frequent direction changes~\cite{Gasparetto.2010}. Furthermore, effectively coordinating the movement of multiple joints to execute rapid direction changes can prove to be a computationally intensive task~\cite{VandeWeghe.2007}. Poor robot configurations can result in prolonged and unnecessary movements, ultimately hampering the overall productivity of the manufacturing process~\cite{Reiter.2016}. %Therefore, advanced control algorithms and motion planning techniques are required to optimize direction changes and minimize transfer time, ensuring efficient and timely execution of manufacturing tasks.
%Minimizing production time is crucial for improving efficiency and throughput.
%Optimal path planning, motion optimization, and parallel processing techniques can be employed to reduce non-value adding movements while leveraging redundant \acrshort{DoF}s effectively, for significant process improvement~\cite{Boscariol.2020}.
If constrained effectively, the redundant \acrshort{DoF}s can serve as helpful tool to reduce the direction changes in the joints of a industrial robot, while its execution of a toolpath. 




Energy use is also a significant concern in manufacturing systems employing redundant \acrshort{DoF}s~\cite{Doan.2016}. The presence of additional joints and their non-optimal usage can require more power to operate, leading to increased energy consumption. As energy efficiency becomes a priority in modern manufacturing, mitigation strategies regarding this problem are necessary~\cite{Boscariol.2020, Boscariol.2019}. One possible solution, is the optimal constraining of the redundant \acrshort{DoF}s which are not directly defined by the toolpath. No method is considering the influence of the redundant \acrshort{DoF}s on energy consumption, while simultaneously optimizing for the process variables mentioned before.
%energy management strategies are necessary to mitigate the increased power demand

%Maximum load capacity is another area influenced by redundant degrees of freedom. The presence of additional joints can introduce structural limitations, affecting the robot's ability to handle heavy loads. Structural analysis and design considerations are crucial to ensure that the robot's mechanical structure can withstand the required load capacity without compromising its integrity or performance. This may involve using high-strength materials, optimizing structural configurations, and employing robust safety mechanisms to prevent overloading.

While redundant \acrshort{DoF}s may introduce potential conflicts and require special attention, they can also significantly enhance performance in manufacturing systems~\cite{Ayten.2016}. The added \acrshort{DoF}s increase flexibility and adaptability, enabling the robot to carry out complex tasks more efficiently. Redundancy enables multiple approaches to achieve a desired end-effector position or orientation. By effectively utilizing the surplus of \acrshort{DoF}s, manufacturing systems can enhance their performance, increase efficiency, and exhibit greater process stability while performing diverse tasks~\cite{Boscariol.2020}. 

Currently, there is no integrated system that can evaluate a computed tool path based on the configuration of the manufacturing machine as well as the chosen process variables, such as combined direction changes or stiffness. %, in conjunction with available \acrshort{CAM} systems.
Additionally, there is no published method that provides an optimal or near-optimal solution for defining the necessary constraints for a specific goal like for example, minimizing energy usage while at the same time reducing joint accelerations.


\section{Objective}\label{OBJECTIVE}
The definition of the constraints for the redundant \acrshort{DoF}s, as mentioned in Chapter \ref{Problem Formulation}, does not affect the relative tool path generated by the \acrshort{CAM} software. As such, a methodical approach to optimize these constraints without altering the toolpath, in terms of energy demand of the machine, accelerations in the joints and further process variables is required. Currently, no literature provides a comprehensive analysis or methodology regarding this global optimization problem.
This work aims to attain a methodical approach that analyzes a set of constraints and evaluates the influence of those constraints on a set of defined process variables. This work is focused on a 6-axis industrial robot with a rotary-tilt table, whereby the results should also be transferable to other machines. Furthermore, the validations is limited to the manufacturing processes of wire arc additive manufacturing (\acrshort{WAAM}) and milling. 


First, the influence of the constraints on relevant process variables (energy demand, joint turnover, speed and acceleration peaks, total joint movements) in a manufacturing process such as \acrshort{WAAM} is assessed. Subsequently, a process evaluation is elaborated, by means of which the process quality can be determined. %Depending on the respective process variables, approximation methods or machine learning methods are investigated for the process evaluation.
The process quality as a one-dimensional variable is determined by weighting the process variables. Subsequently, a method for the optimization of the constraints is elaborated. This task corresponds to an optimization problem in which the process quality is maximized by selecting suitable constraints. 
