% Deckblatt
\chapter*{Scope of Work}

%\markright{Aufgabenstellung} 	% Kolumnentitel manuell auf "Aufgabenstellung"

\textbf{Title of the Master's Thesis:}\\
\Large{Methodical Approach for Analyzing Process Parameters and Optimizing Boundary Conditions in Mutli-Axis Robot Programs}\\
\newline
\normalsize{\textbf{Titel der Master's Thesis:}}\\
\Large{Methodischer Ansatz zur Analyse von Prozessparametern und Optimierung von Randbedingungen in Mutli-Achs-Roboterprogrammen}
\normalsize

\begin{tabbing}
	\hspace{7em} 		\= \hspace{13em}			\= \hspace{7em} 		\= \kill
				\> 						\> \\
	\textbf{Author:} \>  Jan Nalivaika	\> \textbf{Supervisor:} 	\>Ludwig Siebert  \\
	\textbf{Issue:} 	\> 02.10.2023	\> \textbf{Submission:} 	\> 29.03.2024
\end{tabbing}

\vspace{5mm}
\textbf{Motivation:}\\
Computer-aided manufacturing (CAM) is used to automatically generate tool paths for computer numerical control (CNC) machines. The CAM software considers the models of the raw and finished part as well as the constraints of the machine, the tools, and the manufacturing technology. Together with user-configurable parameters, tool paths for 3-axis, 5-axis, and robot-based machine tools are generated. The growing demand for flexibility in machine tools, such as the use of multiple manufacturing technologies in one machine or automated loading and unloading, has led to many machine tools being equipped with additional mechanical axes. Examples include robots mounted on linear axes and rotary-tilt tables.
The tool paths created in CAM programs are usually defined by five degrees of freedom. The first three are the translational axes X, Y, and Z. The tilting and inclining of the tool are defined by the A- and B-axes. Occasionally, an additional rotation of the tool (C-axis) around the Z-axis (e.g., for dragging a swivel knife) is defined. Machines with more degrees of freedom than those limited by the toolpath often need user-defined constraints. These constraints are necessary to fully specify the movements of the machine axes. An example is the alignment of a part using the rotary-tilt table so that the Z-axis of the tool always points in the direction of gravity. This is helpful in processes like fused deposition modeling (FDM) and wire arc additive manufacturing (WAAM). 
It is common practice to set the user-defined constraints based on experience. A preliminary literature review indicates that the configuration of these degrees of freedom has an impact on the energy demand and stability of the process.

\vspace{5mm}
\textbf{Objective:}\\
The definition of these constraints does not affect the relative tool path generated by the CAM software. As such, a methodical approach to optimize these constraints in terms of efficiency, speed, and energy demand of the machine is required. Currently, no literature provides a comprehensive analysis or methodology regarding this global optimization problem.
This work aims to attain a methodical approach that analyzes a set of constraints and evaluates the influence of those constraints on a set of defined process variables. It will focus on a 6-axis robot with a rotary-tilt table, whereby the results should also be transferable to other machines. Furthermore, the experiments and validations will be limited to the manufacturing processes of WAAM and milling. 
First, the influence of the constraints on relevant process variables (energy demand, joint turnover, speed and acceleration peaks, total joint movements) in a manufacturing process such as WAAM will be assessed. Subsequently, a process evaluation will be elaborated in the CAM software, by means of which the process quality can be determined. Depending on the respective process variables, approximation methods or machine learning methods will be investigated for the process evaluation. The process quality as a one-dimensional variable will be determined by weighting the process variables. Subsequently, a method for the optimization of the constraints will be elaborated. This task corresponds to an optimization problem in which the process quality will be maximized by selecting suitable constraints. 

\vspace{5mm}
\textbf{Procedure and working method:}
The following work packages are conducted within this thesis:

\begin{itemize}
\item Literature research
\item Familiarization with WAAM, milling machines, and CAM software
\item Selection of suitable process parameters
\item Elaboration of the proposed method in a suitable programming language
\item Verification and validation of the elaborated method
\item Documentation of the work

	 
\end{itemize}
\vspace{1.0cm}
\textbf{Agreement:}\\

Through the supervision of B.Sc. Jan Nalivaika intellectual property of the \textit{iwb} is incorporated in this work. Publication of the work or transfer to third parties requires the permission of the chair holder. I agree to the archiving of the work in the library of the \textit{iwb}, which is only accessible to \textit{iwb} staff, as inventory and in the digital student research project database of the \textit{iwb} as a PDF document.
\vfill

\IWBaddressCityChair, 29.03.2024
\vspace{2.5cm}\\
\begin{tabular}{p{0.5\linewidth}p{0.5\linewidth} }
	Prof. Dr.-Ing.		& B.Sc.\\
Michael F. Zäh  	& Jan Nalivaika
\end{tabular}
